---
title: "Playmate analysis"
author: "Yevgeny V.Yorkhov"
date: "03/15/2015"
output: html_document
---

# Playboy January 1991 playmate is the model with the most popular body sizes!

## Synopsis
**[Stacy Arthur](http://en.wikipedia.org/wiki/Stacy_Arthur)** has the most popular body measurements (Bust, Cup, Waist, Hips, Height, Weight, BMI, WH) from 1975 till 2008.
  
## The dataset

The data come from the following URL: http://archive.wired.com/wired/archive/17.02/1702_Infoporn_Playmate_Data.xls .

I used the following variables in analysis:

1. Bust - Bust size in inches
2. Cup - Cup size. US and Canada standards. ( [a=1, b=2, c=3, d=4, dd=5, ddd=6, e=7, ee=8](http://en.wikipedia.org/wiki/Brassiere_measurement#United_States_and_Canada) )
3. Waist - Waist size in inches
4. Hips - Hip size in inches
5. Height - in inches
6. Weight - in lb
7. BMI - [Body Mass Index](http://en.wikipedia.org/wiki/Body_mass_index)
8. WH - [Waist-hip ration](http://en.wikipedia.org/wiki/Waist%E2%80%93hip_ratio)

## 1. Data preparation

The dataset is small so I prepared 8 new datasets which contain complete cases of each variable.

```{r, load_and_prepare_data, cache=TRUE}
library("xlsx")
library("lubridate")

## Download data
#url <- "http://archive.wired.com/wired/archive/17.02/1702_Infoporn_Playmate_Data.xls"
#download.file(url, destfile = "1702_Infoporn_Playmate_Data.xls")

## Read data
df <- read.xlsx("1702_Infoporn_Playmate_Data.xls", sheetIndex=1, startRow=1, endRow=663)
names(df) <- c("Month", "Year", "Bust", "Cup", "Waist", "Hips", "Height", "Weight", "BMI", "WH")

## Convert into numeric
for(i in 3:10) {
  df[,i] <- as.numeric(gsub('[A-Z a-z/]', '', df[,i]))
}

## Convert Year and Month into Date type
tmp <- paste(paste(df$Year, df$Month), "01")
df$Date <- as.Date(tmp, format="%Y %B %d")

## Find complete cases, make new data frames
tmp <- subset(df, complete.cases(df$Bust)==TRUE)
df_Bust <- tmp[!(tmp$Bust == "NOT LISTED"),]
df_Bust$Bust <- as.numeric(gsub('[A-Z a-z/]', '', df_Bust$Bust))

tmp <- subset(df, complete.cases(df$Cup)==TRUE)
df_Cup <- tmp[!(tmp$Cup == "NOT LISTED"),]

tmp <- subset(df, complete.cases(df$Waist)==TRUE)
df_Waist <- tmp[!(tmp$Waist == "NOT LISTED"),]
df_Waist$Waist <- as.numeric(gsub('[A-Z a-z/]', '', df_Waist$Waist))

tmp <- subset(df, complete.cases(df$Hips)==TRUE)
df_Hips <- tmp[!(tmp$Hips == "NOT LISTED"),]
df_Hips$Hips <- as.numeric(gsub('[A-Z a-z/]', '', df_Hips$Hips))

tmp <- subset(df, complete.cases(df$Height)==TRUE)
df_Height <- tmp[!(tmp$Height == "NOT LISTED"),]
df_Height$Height <- as.numeric(gsub('[A-Z a-z/]', '', df_Height$Height))

tmp <- subset(df, complete.cases(df$Weight)==TRUE)
df_Weight <- tmp[!(tmp$Weight == "NOT LISTED"),]
df_Weight$Weight <- as.numeric(gsub('[A-Z a-z/]', '', df_Weight$Weight))

tmp <- subset(df, complete.cases(df$BMI)==TRUE)
df_BMI <- tmp[!(tmp$BMI == 0),]

tmp <- subset(df, complete.cases(df$WH)==TRUE)
df_WH <- tmp[!(tmp$WH == 0),]
```

Let's look at the new datasets
```{r, look_at_unique}
unique(df_Bust$Bust)
mm_bust <- c(mean(df_Bust$Bust), median(df_Bust$Bust))
mm_bust

unique(df_Cup$Cup)
mm_cup <- c(mean(df_Cup$Cup), median(df_Cup$Cup))
mm_cup

unique(df_Waist$Waist)
mm_waist <- c(mean(df_Waist$Waist), median(df_Waist$Waist))

unique(df_Hips$Hips)
mm_hips <- c(mean(df_Hips$Hips), median(df_Hips$Hips))
mm_hips

unique(df_Height$Height)
mm_height <- c(mean(df_Height$Height), median(df_Height$Height))
mm_height

unique(df_Weight$Weight)
mm_weight <- c(mean(df_Weight$Weight), median(df_Weight$Weight))
mm_weight

mm_bmi <- c(mean(df_BMI$BMI), median(df_BMI$BMI))
mm_bmi

mm_wh <- c(mean(df_WH$WH), median(df_WH$WH))
mm_wh
```
Looks good. Let's continue with some plotting.

### Plot the frequencies of body measurements.

```{r, plot_sizes, echo=FALSE}
par(mfrow=c(2,2))
## Plot Bust size
hist(df_Bust$Bust, main="The frequency of\nBust size", xlab="Bust", col="cyan")
abline(v=mean(df_Bust$Bust), lty=2, lwd=3, col="red")
abline(v=median(df_Bust$Bust), lty=2, lwd=3, col="blue")
text(mean(df_Bust$Bust), 100, labels = paste("mean = ", as.character(round(mean(df_Bust$Bust), digits=4))),
     pos = 4, col="red")
text(mean(df_Bust$Bust), 120, labels = paste("median = ", as.character(round(median(df_Bust$Bust), digits=4))),
     pos = 4, col="blue")

## Plot Cup size
hist(df_Cup$Cup, main="The frequency of\n Cup size", xlab="Cup", col="cyan")
abline(v=mean(df_Cup$Cup), lty=2, lwd=3, col="red")
abline(v=median(df_Cup$Cup), lty=2, lwd=3, col="blue")
text(mean(df_Cup$Cup), 50, labels = paste("mean = ", as.character(round(mean(df_Cup$Cup), digits=4))), 
     pos = 4, col="red")
text(mean(df_Cup$Cup), 60, labels = paste("median = ", as.character(round(median(df_Cup$Cup), digits=4))), 
     pos = 4, col="blue")

## Plot Waist size
hist(df_Waist$Waist, main="The frequency of\n Waist size", xlab="Waist", col="cyan")
abline(v=mean(df_Waist$Waist), lty=2, lwd=3, col="red")
abline(v=median(df_Waist$Waist), lty=2, lwd=3, col="blue")
text(mean(df_Waist$Waist), 200, labels = paste("mean = ", as.character(round(mean(df_Waist$Waist), digits=4))), 
     pos = 4, col="red")
text(mean(df_Waist$Waist), 230, labels = paste("median = ", as.character(round(median(df_Waist$Waist), digits=4))), 
     pos = 4, col="blue")

## Plot Hips size
hist(df_Hips$Hips, main="The frequency of Hips size", xlab="Hips", col="cyan")
abline(v=mean(df_Hips$Hips), lty=2, lwd=3, col="red")
abline(v=median(df_Hips$Hips), lty=2, lwd=3, col="blue")
text(mean(df_Hips$Hips), 150, labels = paste("mean = ", as.character(round(mean(df_Hips$Hips), digits=4))), 
     pos = 3, col="red")
text(mean(df_Hips$Hips), 170, labels = paste("median = ", as.character(round(median(df_Hips$Hips), digits=4))), 
     pos = 3, col="blue")

par(mfrow=c(2,2))
## Plot Height size
hist(df_Height$Height, main="The frequency of Height", xlab="Height", col="cyan")
abline(v=mean(df_Height$Height), lty=2, lwd=3, col="red")
abline(v=median(df_Height$Height), lty=2, lwd=3, col="blue")
text(mean(df_Height$Height), 60, labels = paste("mean = ", as.character(round(mean(df_Height$Height), digits=4))),
     pos = 4, col="red")
text(mean(df_Height$Height), 80, labels = paste("median = ", as.character(round(median(df_Height$Height), digits=4))),
     pos = 4, col="blue")

## Plot Weight size
hist(df_Weight$Weight, main="The frequency of Weight", xlab="Weight", col="cyan")
abline(v=mean(df_Weight$Weight), lty=2, lwd=3, col="red")
abline(v=median(df_Weight$Weight), lty=2, lwd=3, col="blue")
text(mean(df_Weight$Weight), 100, labels = paste("mean = ", as.character(round(mean(df_Weight$Weight), digits=4))),
     pos = 4, col="red")
text(mean(df_Weight$Weight), 120, labels = paste("median = ", as.character(round(median(df_Weight$Weight), digits=4))),
     pos = 4, col="blue")

## Plot BMI size
hist(df_BMI$BMI, main="The frequency of BMI", xlab="BMI", col="cyan") 
abline(v=mean(df_BMI$BMI), lty=2, lwd=3, col="red")
abline(v=median(df_BMI$BMI), lty=2, lwd=3, col="blue")
text(mean(df_BMI$BMI), 60, labels = paste("mean = ", as.character(round(mean(df_BMI$BMI), digits=4))),
     pos = 4, col="red")
text(mean(df_BMI$BMI), 80, labels = paste("median = ", as.character(round(median(df_BMI$BMI), digits=4))),
     pos = 4, col="blue")

## Plot WH size
hist(df_WH$WH, main="The frequency of Waist/Hips", xlab="WH", col="cyan")
abline(v=mean(df_WH$WH), lty=2, lwd=3, col="red")
abline(v=median(df_WH$WH), lty=2, lwd=3, col="blue")
text(mean(df_WH$WH), 200, labels = paste("mean = ", as.character(round(mean(df_WH$WH), digits=4))),
     pos = 4, col="red")
text(mean(df_WH$WH), 220, labels = paste("median = ", as.character(round(median(df_WH$WH), digits=4))),
     pos = 4, col="blue")

```

### Fix errors in dataset
Hips size frequency looks strange. Let's dig into this.

List models whos hips size < 30''
```{r, fix_dataset_hips_err}
df_Hips[df_Hips$Hips < 30, c(1,2,3,4,5,6,7,8,9,10)]
```

Let's compare the hips size data of these models with data from Wikipedia.

1. Feb 1991 - Model name **[Cristy Thom](http://en.wikipedia.org/wiki/List_of_Playboy_Playmates_of_1991)**, real hips size: 36" - **incorrect**, need to fix in the dataset.
2. Apr 2003 - Model name **[Carmella DeCesare](http://en.wikipedia.org/wiki/Carmella_DeCesare)**, real hips size: 27 - **correct**
3. Jan 2005 - Modle name **[Destiny Davis](http://en.wikipedia.org/wiki/List_of_Playboy_Playmates_of_2005)**, real hips size: 34" - **incorrect**, need to fix in the dataset.

```{r, fix_hips_data}
## Fix Feb 1991
df_Hips[df_Hips$Hips == 26, ]$Hips <- 36

## Fix Jan 2005
df_Hips[df_Hips$Hips == 24, ]$Hips <- 34

unique(df_Hips$Hips)
mm_hips <- c(mean(df_Hips$Hips), median(df_Hips$Hips))
mm_hips
```


## 2. Data processing
Find the model who has the most popular sizes of body measurement. 

### Find the models who's sizes are the closest to mean values.
```{r, find_by_mean}
prec <- 1
bust <- round(mean(df_Bust$Bust), digits = prec)
cup <- round(mean(df_Cup$Cup), digits = prec)
waist <- round(mean(df_Waist$Waist), digits = prec)
hips <- round(mean(df_Hips$Hips), digits = prec)
height <- round(mean(df_Height$Height), digits = prec)
weight <- round(mean(df_Weight$Weight), digits = prec)
bmi <- round(mean(df_BMI$BMI), digits = prec)
wh <- round(mean(df_WH$WH), digits = prec)

df1 <- subset(df, complete.cases(df)==TRUE)
range <- 1

df1[ 
  df1$Bust < bust+range & df1$Bust > bust-range &
  df1$Cup < cup+range & df1$Cup > cup-range &
  df1$Waist < waist+range & df1$Waist > waist-range &
  df1$Hips < hips+range & df1$Hips > hips-range &
  df1$Height < height+range & df1$Height > height-range &
  df1$Weight < weight+range & df1$Weight > weight-range &
  df1$BMI < bmi+range & df1$BMI > bmi-range &
  df1$WH < wh+range & df1$WH > wh-range 
, c(1,2,3,4,5,6,7,8,9,10)]
```

The model is **[Stacy Arthur](http://en.wikipedia.org/wiki/Stacy_Arthur)**

### Find the models who's sizes are the closest to median values.
```{r, find_by_median}
bust <- round(median(df_Bust$Bust), digits = prec)
cup <- round(median(df_Cup$Cup), digits = prec)
waist <- round(median(df_Waist$Waist), digits = prec)
hips <- round(median(df_Hips$Hips), digits = prec)
height <- round(median(df_Height$Height), digits = prec)
weight <- round(median(df_Weight$Weight), digits = prec)
bmi <- round(median(df_BMI$BMI), digits = prec)
wh <- round(median(df_WH$WH), digits = prec)

range <- 1.001

df1[ 
    df1$Bust < bust+range & df1$Bust > bust-range &
    df1$Cup < cup+range & df1$Cup > cup-range &
    df1$Waist < waist+range & df1$Waist > waist-range &
    df1$Hips < hips+range & df1$Hips > hips-range &
    df1$Height < height+range & df1$Height > height-range &
    df1$Weight < weight+range & df1$Weight > weight-range &
    df1$BMI < bmi+range & df1$BMI > bmi-range &
    df1$WH < wh+range & df1$WH > wh-range 
  , c(1,2,3,4,5,6,7,8,9,10)]
```

## 3. Result
Here are 9 models.

1. Jan 1992 - **[Stacy Arthur](http://www.boobpedia.com/boobs/Stacy_Arthur)**
2. March 1992 - **[Tylyn John](http://www.boobpedia.com/boobs/Tylyn_John)**
3. March 1997 - **[Jennifer Miriam](http://www.boobpedia.com/boobs/Jennifer_Miriam)**
4. June 1997 - **[Carrie Stevens](http://www.boobpedia.com/boobs/Carrie_Stevens)**
5. December 2004 - **[Tiffany Fallon](http://www.boobpedia.com/boobs/Tiffany_Fallon)**
6. March 2006 - **[Monica Leigh](http://www.boobpedia.com/boobs/Monica_Leigh)**
7. July 2007 - **[Tiffany Selby](http://www.boobpedia.com/boobs/Tiffany_Selby)**
8. September 2008 - **[Valerie Mason](http://www.boobpedia.com/boobs/Valerie_Mason)**
9. October 2008 - **[Kelly Carrington](http://www.boobpedia.com/boobs/Kelly_Carrington)**
